---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# dimmodelR

<!-- badges: start -->
<!-- badges: end -->

Business-intelligence tools such as Power BI, Tableau and Qlik work best when 
they consume data organized as a [dimensional model](https://en.wikipedia.org/wiki/Dimensional_modeling).
A common task before using these tools is therefore to transform data according 
to this model's design principles. This is also know as [star schema](https://en.wikipedia.org/wiki/Star_schema) 
design.

The `dimmodelR` package aims to make this easy using R. It provides a set of 
functions for automating the creation and refreshment of a central data model 
consisting of multiple [fact tables](https://en.wikipedia.org/wiki/Fact_table) 
referencing a set of shared [dimensions](https://en.wikipedia.org/wiki/Dimension_(data_warehouse)#Dimension_table).


## Installation

You can install the development version of dimmodelR from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("cvermehren/dimmodelR")
```

## Workflow

The following workflow is supported:

1. Define a star schema from one or more data frames
2. Populate the schema with data and set up incremental refresh
4. Save the model as csv or parquet files and serve them to a BI tool


## Define a star schema

A star schema can be created from multiple data frames, but needs to be built 
gradually. We will start with the included demo dataset `campaign_metrics`:

```{r, eval=T}
library(dimmodelR)

data(campaign_metrics)

str(campaign_metrics)
```

Typically fact tables are formed by numerical columns, while dimension columns
are character type.

To define a star schema we only need to specify the dimensions. Fact tables are 
added later when populating the model with data.

```{r, eval=T}
# Define dimensions
dimensions = list(dim_channel = c("source", "medium", "campaign"))

# Initiate the model
dm <- dm_model(campaign_metrics, dimensions)

str(dm)
```

The function `dm_model()` has extracted one dimension, `dim_channel`, from the 
data frame and added a [surrogate key](https://en.wikipedia.org/wiki/Surrogate_key) 
column named `channel_key`. 

The surrogate key will function as the [primary key](https://en.wikipedia.org/wiki/Primary_key) 
of `dim_channel` and will be inserted into the fact table as a [foreign key](https://en.wikipedia.org/wiki/Foreign_key) 
when we populate the model with data.

Before we do so, let's extend the model with a few more dimensions using the 
`web_metrics` and `email_metrics` datasets.

```{r, eval=T}
data("web_metrics")

# Inspect the dataset
str(web_metrics)

```

The `web_metrics` dataset introduces two new dimension columns: `view_name` and 
`country`. These can be added to the model as long as we remember to use the 
model itself (`dm`) as an argument in the function.

```{r, eval=T}
dimensions = list(
  dim_channel = c("source", "medium", "campaign"),
  dim_market = c("view_name", "country")
)

# Extend the model using the model itself (`dm`) as an argument
dm <- dm_model(web_metrics, dimensions, dm)

str(dm)
```

The model now consist of two dimensions. Let's add one more using the final demo 
dataset.

```{r, eval=T}
data("email_metrics")

dimensions = list(dim_email = c("email"))

dm <- dm_model(email_metrics, dimensions, dm)

str(dm)
```


## Populate the model with data

We now have a dimensional model consisting of three dimensions, but no facts. To
populate the model with facts, you only need to pass the model and a named list 
of source data to `dm_refresh()`. The function will automatically match the 
source data with dimensions in the model and extract fact tables accordingly.


```{r, eval=T}
# Make a named list of source data from which fact tables should be extracted
facts <- list(
  fct_email = email_metrics, 
  fct_campaign = campaign_metrics,
  fct_web = web_metrics
)

# Populate the model with data
dm <- dm_refresh(dm, facts)

str(dm)
```
The model now contains all data from the three data frames (source data) 
organized as a dimensional model. 

The refresh function has added not only three fact tables, but also new rows to 
the dimensions. The initial model is based only on a sample of source data. 
When adding the entire datasets using the refresh function, the dimensions are 
updated with rows that are new to the model.

The function also handles shared dimension. Two of the fact tables 
(`fct_campaign` and `fct_web`) share the `dim_channel` dimension. In other words, 
the function inserts this dimension's primary key (`channel_key`) into both of 
these fact tables. It automatically identifies how to do so by comparing source 
data and existing dimensions in the model.

To save the result as csv files:

```{r, eval=T}
dm_write_csv(dm, "my-folder")
 
list.files(path = "my-folder", recursive = TRUE)

```

The `dm_write_csv` function saves the result in two sub-directories `dimensions` 
and `fact_tables`. It will always overwrite the files and, therefore, is not 
meant for incremental refresh.


## Incremental refresh

To refresh your model incrementally, simply feed the model with the newest 
source data followed by `dm_write_parquet`.

Let's simulate how this would work.

```{r, eval=T}

# Load demo data
data("web_metrics")
data("campaign_metrics")
data("email_metrics")

# # Ad partitioning columns
# web_metrics$month <- lubridate::month(web_metrics$date)
# campaign_metrics$month <- lubridate::month(campaign_metrics$date)
# email_metrics$month <- lubridate::month(email_metrics$date)
# 
# # Create campaig_metrics model
# dimensions <- list(dim_channel = c("source", "medium", "campaign"))
# dm <- dm_model(campaign_metrics, dimensions)
# 
# # Extend model with web_metrics
# dimensions <- list(
#   dim_channel = c("source", "medium", "campaign"), 
#   dim_market = c("view_name", "country"))
# dm <- dm_model(web_metrics, dimensions, dm)
# 
# # Extend model with email_metrics
# dimensions <- list(dim_channel = c("email"))
# dm <- dm_model(email_metrics, dimensions, dm)

# Make looping-list for incremental refresh
months <- sort(unique(web_metrics$month))

# Simulate incremental refresh
for (i in months) {
  new_campaign_metrics <- campaign_metrics[which(campaign_metrics$month == i),]
  new_web_metrics <- web_metrics[which(web_metrics$month == i),]
  new_email_metrics <- email_metrics[which(email_metrics$month == i),]
  
  facts <- list(
    fct_campaign = new_campaign_metrics,
    fct_web = new_web_metrics,
    fct_email = new_email_metrics
    )
  
  dm <- dm_refresh(dm, facts)
  
  dm_write_parquet(dm,path = "my-incremental-model", partitioning = c("year", "month"))
}

# Viewing the created files for fct_campaign
list.files(
  path = "my-incremental-model/fact_tables/fct_campaign", 
  recursive = TRUE
  )
```

