---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dimmodelR

<!-- badges: start -->
<!-- badges: end -->

Transforming source data to a [dimensional model](https://en.wikipedia.org/wiki/Dimensional_modeling) 
is one of the most difficult, time-consuming and error-prone tasks within data 
engineering and business intelligence. 

`dimmodelR` aims to make this task  easy. It provides a set of R functions which 
automate the process of creating and refreshing a central star-schema model 
consisting of multiple fact tables and shared dimension tables. 


## Features

* Transform a data frame into a [star schema](https://en.wikipedia.org/wiki/Star_schema) 
  with dimensions and fact tables related by [surrogate keys](https://en.wikipedia.org/wiki/Surrogate_key)
* Extend a star schema with more transformations turning it into a central model also known as 
  [fact constellation](https://en.wikipedia.org/wiki/Fact_constellation)
* Easily refresh a central model incrementally as new source data become available
* Save the model as csv or parquet files in a practical predefined folder structure


## Installation

You can install the development version of dimmodelR from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("cvermehren/dimmodelR")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r, eval=FALSE}
library(dimmodelR)
## basic example code
```

